---
title: "RWC Semi Final: NZ v ENG"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: bootstrap
    css: static/css/custom-dash.css
---

<!-- Custom JS to help with formatting non-standard content -->
<script src="static/js/custom-flexdashboard-components.js"></script> 

```{r "setup", include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.align = 'center')
library(flexdashboard)
library(htmltools)
library(data.table)
library(plotly)
library(glue)
library(gt)
library(paletteer)
library(emo)
library(kableExtra)
source('../r-code/rugby_plots.R')
```


```{r "data-source", include=FALSE}
### SET THESE VALUES TO CUSTOMISE
logos_pth <- 'static/imgs/team-colours'
score_events <- c("T5", "C2", "P3")
posix_KO <- as.POSIXct("2019-10-26 07:30:00", tz="UTC")
posix_FT <- as.POSIXct("2019-10-26 10:30:00", tz="UTC")
ref_regex <- '( ref|Owens|owens|Nige |nige |Nigel|nigel)'
match_data <- jsonlite::fromJSON('../data/reddit-match-threads/dn64ve_matchdata.json')
mt_comms <- fread('../data/reddit-match-threads/dn64ve_final.csv', encoding = 'UTF-8')
timeline <- fread('../data/reddit-match-threads/dn64ve_timeline.csv')
TEAM_0_NAME <- tolower(match_data$teams[1, 'name'])
TEAM_1_NAME <- tolower(match_data$teams[2, 'name'])
TEAM_0_ID <- match_data$teams[1, 'id']
TEAM_1_ID <- match_data$teams[2, 'id']
```


```{r "data-prep"}
# Add R POSIXct type datetime columns
mt_comms[,posix_dt := as.POSIXct(created_utc, origin='1970-01-01', tz='UTC')]
timeline[,posix_dt := as.POSIXct((millis/1000)+(gmt_offset*60*60), origin='1970-01-01', tz='UTC')]
timeline[,posix_dt := as.POSIXct(unlist(lapply(posix_dt, wr_timestamp_bst)), origin='1970-01-01', tz='UTC')]
timeline[is.na(team_id), event_team := 'neutral']

# Clean up mt_comms
mt_comms[flair_country=='', flair_country := 'unknown']

# Clean up the timeline info
timeline[team_id == TEAM_0_ID, event_team:=TEAM_0_NAME]
timeline[team_id == TEAM_1_ID, event_team:=TEAM_1_NAME]
timeline[is.na(event_team), event_team:='']
timeline[,graph_lab := glue_data(.SD, '{sprintf("%02d", match_time%/%60)}:{sprintf("%02d", match_time%%60)}<br>{label} {titleize(event_team)}<br>{score0} - {score1}')]

# Merge timeline into comment data
combodt <- rbindlist(list(mt_comms, timeline), fill = T)
setorder(combodt, posix_dt)
combodt[is.na(score_praw), score_praw:=0]
combodt[is.na(score), score:=0]
combodt[is.na(vader_score), vader_score:=0]
combodt[is.na(google_score), google_score:=0]
combodt[is.na(google_magnitude), google_magnitude:=0]
combodt[,allegiance := ifelse(flair_country %in% c(TEAM_0_NAME, TEAM_1_NAME), tolower(flair_country), 'neutral')]
combodt[,ref:=grepl(ref_regex, plaintext)]
combodt[, c("cum_google", "cum_vader") := .(cumsum(google_score), cumsum(vader_score))]
combodt[, c("cum_google_ref", "cum_vader_ref") := .(cumsum(google_score), cumsum(vader_score)), by=ref]
combodt[, c("pre_2", "pre_5", "end") := .(posix_dt - 2*60, posix_dt - 5*60, posix_dt)]

# Add rolling mean columns
combodt[, c("rm_2mins_google", "rm_2mins_vader") := 
          .SD[.SD, on=.(posix_dt>=pre_2, posix_dt<=end), 
              by=.EACHI, lapply(.SD, mean), .SDcols=c("google_score", "vader_score")
              ][,(1L:2L) := NULL]
]
combodt[, c("rm_2mins_google_flair", "rm_2mins_vader_flair") := 
          .SD[.SD, on=.(posix_dt>=pre_2, posix_dt<=end), 
              by=.EACHI, lapply(.SD, mean), .SDcols=c("google_score", "vader_score")
              ][,(1L:2L) := NULL],
        by=allegiance
]
combodt[, c("rm_2mins_google_ref", "rm_2mins_vader_ref") := 
          .SD[.SD, on=.(posix_dt>=pre_2, posix_dt<=end), 
              by=.EACHI, lapply(.SD, mean), .SDcols=c("google_score", "vader_score")
              ][,(1L:2L) := NULL],
        by=ref
]
combodt[, c("rm_5mins_google", "rm_5mins_vader") := 
          .SD[.SD, on=.(posix_dt>=pre_5, posix_dt<=end), 
              by=.EACHI, lapply(.SD, mean), .SDcols=c("google_score", "vader_score")
              ][,(1L:2L) := NULL]
]
combodt[, c("rm_5mins_google_flair", "rm_5mins_vader_flair") := 
          .SD[.SD, on=.(posix_dt>=pre_5, posix_dt<=end), 
              by=.EACHI, lapply(.SD, mean), .SDcols=c("google_score", "vader_score")
              ][,(1L:2L) := NULL],
        by=allegiance
]
combodt[, c("rm_5mins_google_ref", "rm_5mins_vader_ref") := 
          .SD[.SD, on=.(posix_dt>=pre_5, posix_dt<=end), 
              by=.EACHI, lapply(.SD, mean), .SDcols=c("google_score", "vader_score")
              ][,(1L:2L) := NULL],
        by=ref
]

# Add team colors
team_cols <- fread(paste(logos_pth, 'team-colours.csv', sep='/'))
timeline[,event_bg_col:=sapply(event_team, get_team_bg_col, team_cols=team_cols)]
timeline[,event_text_col:=sapply(event_team, get_team_text_col, team_cols=team_cols)]
timeline[,event_border_col:=sapply(event_team, get_team_border_col, team_cols=team_cols)]
score_data <- timeline[event %in% score_events]
```


```{r "helper-funcs", include=FALSE}

get_sentiment_icon <- function(x){
  dplyr::case_when(
    x > 0.25 ~ "- far fa-smile", 
    x < -0.25 ~ "- far fa-frown", 
    TRUE ~ "- far fa-meh"
  )
}

get_sentiment_col <- function(x){
  dplyr::case_when(
    x > 0.25 ~ "success", 
    x < -0.25 ~ "danger", 
    TRUE ~ "warning"
  )
}

get_score_col <- function(x){
  dplyr::case_when(
    x > 0 ~ "success", 
    x < 0 ~ "danger", 
    TRUE ~ "warning"
  )
}

get_score_icon <- function(x){
  dplyr::case_when(
    x > 0 ~ "far fa-arrow-alt-circle-up", 
    x < 0 ~ "far fa-arrow-alt-circle-down", 
    TRUE ~ "far fa-minus-square"
  )
}

get_comment_context <- function(comment, all_comments) {
  context <- vector("list", length=(comment$depth+1))
  context[[1]] <- comment
  last_comm <- comment
  i <- 2
  while(TRUE){
    pids <- strsplit(last_comm$parent_id, '_')[[1]]
    if(pids[1] == 't1'){
      last_comm <- as.list(all_comments[id==pids[2],.(id, author, plaintext, score_praw, parent_id, depth)]) 
    }
    else{break}
    context[[i]] <- last_comm
    i <- i+1
  }
  return(Filter(Negate(is.null), rev(context)))
}

get_reddit_link <- function(permalink) {
  return(glue("https://reddit.com{permalink}"))
}

reddit_comment <- function(comment, sentiment = F) {
  if(sentiment) {
    score <- as.numeric(sigfig(comment$google_score))
    score_status <- get_score_col(score)
    score_icon <- get_sentiment_icon(score)
  }
  else {
    score <- comment$score_praw
    score_status <- get_score_col(score)
    score_icon <- get_score_icon(score)
  }
  return (
    div(class = "reddit-comment",
     div(class = glue("reddit-comment-meta-outer bg-{score_status}"),
      div(class = "reddit-comment-meta-inner",
        p(class = "reddit-comment-icon", 
          span(class = glue("{score_icon} text-{score_status}"))
        ),
        p(class = "reddit-comment-score", tags$b(score))
      )
     ),
     div(class = "reddit-comment-content",
       p(class = "reddit-comment-author", 
         tags$b(comment$author), 
         a(class = "reddit-comment-link", 
           href = get_reddit_link(comment$permalink), 
           title = "View on Reddit", 
           span(class = "fab fa-reddit")
          )
        ), 
       p(class = "reddit-comment-body", comment$plaintext)
     )
   )
  )
}

```

Row
-----------------------------------------------------------------------

### Match Facts {.no-title}
```{r "Match Facts"}
div(
  class = "match-facts",
  p(class = "match-facts-datetime", 
    format(as.POSIXct((match_data$time$millis/1000), origin = "1970-01-01", tz='UTC'), "%d %b %Y | %H:%M:%S %Z")
  ),
  span(class = "match-facts-header",
    span(class = "match-facts-team0",
      glue('{match_data$teams[1, "name"]}'), 
      img(
        src=get_team_flag(tolower(TEAM_0_NAME), team_cols=team_cols, path=logos_pth), 
        title=glue("{titleize(TEAM_0_NAME)}"), class="flag match-facts-flag"
      ),
    ),
    span(class = "match-facts-v", 'v'),
    span(class = "match-facts-team1",
      img(
        src=get_team_flag(tolower(TEAM_1_NAME), team_cols=team_cols, path=logos_pth), 
        title=glue("{titleize(TEAM_1_NAME)}"), class="flag match-facts-flag"
      ),
      glue('{match_data$teams[2, "name"]}')
    )
  ),
  p(tags$b(glue('{match_data$venue$name}')), br(), glue("{match_data$venue$city}, {match_data$venue$country}")),
  p(glue("{match_data$scores[1]} - {match_data$scores[2]}"), style=css(font_size="3em"))
)
```

Row
-----------------------------------------------------------------------

### Comments
```{r}
ncomments <- nrow(mt_comms) 
valueBox(ncomments, icon = "fa-comments")
```

### Commenters
```{r}
nauthors <- nrow(mt_comms[author!='[deleted]', .N,by=author]) 
valueBox(nauthors, icon = "fa-users")
```

### Average Sentiment
```{r}
avg_sent <- mt_comms[,mean(google_score)] 
valueBox(
  sigfig(avg_sent, 2), 
  icon = get_sentiment_icon(avg_sent),
  color = get_sentiment_col(avg_sent)
)
```

Row
-----------------------------------------------------------------------
### Comment Flow {.no-title .plotly-chart data-width=2}

```{r "comment-flow"}
comment_hist <- plot_ly() %>% 
  add_trace(
      data = combodt, 
      x = ~posix_dt, 
      xbins = list(size = 60000),
      type = 'histogram',
      hoverinfo = 'y',
      marker = list(
        color = get_team_border_col('neutral', team_cols),
        line = list(color = get_team_bg_col('neutral', team_cols), width = 1)
      )
  ) %>% 
  add_trace(
      data = score_data,
      type = 'scatter',
      mode = 'markers',
      marker = list(
        symbol = 'diamond', 
        color = ~event_bg_col, 
        size = 10, 
        line = list(color = ~event_border_col, width=2)
      ),
      x = ~posix_dt, 
      y = 0,
      text = ~graph_lab,
      hoverinfo = "text",
      hoverlabel = list(font = list(color = ~event_text_col), bordercolor = ~event_border_col)
  ) %>%
  layout(
    showlegend = F,
    margin = list(pad = 20),
    title = list(text = "Comments Submitted Every Minute", yanchor = "bottom"),
    xaxis = list(title = "Time", range = c(posix_KO, posix_FT)),
    yaxis = list(title = "Comments", fixedrange = T),
    modebar = list(orientation = 'v'),
    updatemenus = list(
      list(
        type = "buttons",
        y = -0.2,
        x= 0.5,
        xanchor = 'center',
        direction = 'left',
        buttons = list(
          list(method = "restyle",
               args = list("cumulative.enabled", FALSE),
               label = "Standard"), 
          list(method = "restyle",
               args = list("cumulative.enabled", TRUE),
               label = "Cumulative")
        )
      )
    )
  ) %>% 
  config(
    modeBarButtonsToRemove = c(
      'hoverCompareCartesian', 
      'hoverClosestCartesian', 
      'lasso2d', 
      'select2d', 
      'toggleSpikelines'
    )
  )

comment_hist
```

### Where In The World {.no-title data-width=1}

```{r "where-in-the-world"}
mt_comms[,
  .(
    users=length(unique(author)), 
    tot_comments=.N, 
    flag=glue_data(.SD, '<img src="{sapply(tolower(flair_country), get_team_flag, team_cols=team_cols, path=logos_pth)}" title="{titleize(flair_country)}" class="flag table-flag">')),
  by=flair_country
][order(-tot_comments)] %>% 
  gt(id='user-by-flair-tab') %>%
  text_transform(
    locations = cells_body(vars(flair_country)),
    fn = titleize
  ) %>% 
  fmt_passthrough(
    columns = vars(flag),
    escape = F
  ) %>% 
  tab_header(
    title = md("**Where In The World?**"),
    subtitle = md("Users and comments by flair")
  ) %>% 
  cols_label(
    flag = "",
    flair_country = "",
    users = "Users",
    tot_comments = "Total Comments"
  ) %>% 
  cols_move_to_start(
    columns = vars(flag)
  ) %>% 
  cols_align(
    align = 'right',
    columns = vars(flag)
  ) %>% 
  tab_options(
    # Remove the top border from the table
    table.border.top.width = 0,
    container.height = pct(100)
  )
```

Row
-----------------------------------------------------------------------

### Most Positive {.flex-chart}
```{r "positive"}
reddit_comment(as.list(mt_comms[order(-google_score, google_magnitude)][1]), sentiment = T)
```

### Most Negative {.flex-chart}
```{r "negative"}
reddit_comment(as.list(mt_comms[order(google_score, google_magnitude)][1]), sentiment = T)
```

### Highest Score {.flex-chart}
```{r "high"}
reddit_comment(as.list(mt_comms[which.max(score_praw),][1]))
```

### Lowest Score {.flex-chart}
```{r "low"}
reddit_comment(as.list(mt_comms[which.min(score_praw),][1]))
```


Row {data-height=700}
-----------------------------------------------------------------------

### Sentiment: 2 Minute Rolling Average {.no-title .plotly-chart}

```{r "sent-plot"}
flair_match_data <- combodt[
  posix_dt > posix_KO & 
  posix_dt < posix_FT,
  .(posix_dt, rm_2mins_google_flair, allegiance)
]

rolling_sent_plt <- function(TEAM, comment_data, score_data, team_cols) {
  plot_ly() %>% 
    add_trace(
      data = comment_data[allegiance==TEAM], 
      x = ~posix_dt, 
      y = ~rm_2mins_google_flair, 
      type = 'scatter',
      mode = 'lines', 
      fill = "tozeroy", 
      line = list(color=hex_to_rgba(get_team_colour_no_white(TEAM, team_cols))),
      fillcolor = hex_to_rgba(get_team_colour_no_white(TEAM, team_cols), 0.5), 
      hoverinfo = "y"
  ) %>% 
    add_trace(
      data = score_data,
      type = 'scatter',
      mode = 'markers',
      marker = list(
        symbol = 'diamond', 
        color = ~event_bg_col, 
        size = 10, 
        line = list(color = ~event_border_col, width=2)
      ),
      x = ~posix_dt, 
      y = 0,
      text = ~graph_lab,
      hoverinfo = "text",
      hoverlabel = list(font = list(color = ~event_text_col), bordercolor = ~event_border_col)
  ) %>% 
    layout(
      yaxis = list(range = c(-1, 1)),
      annotations = list(
        x = 1,
        showarrow = F,
        text = titleize(TEAM),
        xref = "paper",
        yref = "paper",
        xanchor = "left",
        yanchor = "middle",
        y = 0.5,
        textangle = 90,
        font = list(
          color = get_team_text_col(TEAM, team_cols)
        ),
        bgcolor = get_team_bg_col(TEAM, team_cols),
        bordercolor = get_team_border_col(TEAM, team_cols)
      ),
      images = list(
        source = get_team_logo(TEAM, team_cols, logos_pth),
        xref = 'paper',
        yref = 'paper',
        x = 0,
        y = 0.95,
        xanchor = 'middle',
        yanchor = 'middle',
        sizex = 0.3,
        sizey = 0.3
      )
    ) %>% 
  config(
    modeBarButtonsToRemove = c(
      'hoverCompareCartesian', 
      'hoverClosestCartesian', 
      'lasso2d', 
      'select2d', 
      'toggleSpikelines'
    )
  )
}

TEAM_0_NAME_plt <- rolling_sent_plt(TEAM_0_NAME, flair_match_data, score_data, team_cols)

TEAM_1_NAME_plt <- rolling_sent_plt(TEAM_1_NAME, flair_match_data, score_data, team_cols)

neutral_plt <- rolling_sent_plt('neutral', flair_match_data, score_data, team_cols)

subplot(
  TEAM_0_NAME_plt, 
  neutral_plt, 
  TEAM_1_NAME_plt, 
  nrows = 3, 
  shareX = T
) %>% 
  layout(
    title = list(text = "Fanbase Sentiment: 2 Minute Rolling Average", yanchor = "bottom"),
    xaxis = list(title = "Time"),
    yaxis2 = list(matches="y", title = "Average Sentiment"), 
    yaxis3 = list(matches="y"),
    showlegend = FALSE,
    margin = list(r = 30),
    modebar = list(orientation = 'v')
  ) %>% 
  config(
    responsive = TRUE,
    modeBarButtonsToRemove = c(
      'hoverCompareCartesian', 
      'hoverClosestCartesian', 
      'lasso2d', 
      'select2d', 
      'toggleSpikelines'
    )
  )
```

Row
-----------------------------------------------------------------------
### Positive Pauls {.no-title}
```{r "positive-pauls"}
pos_paul <- mt_comms[,
  .(tot_comments=.N, avg_sent=mean(google_score))
  ,by=author
][tot_comments>10][order(-avg_sent)][1:10] %>% 
  gt(rowname_col="author", id='pos-pauls-tab') %>% 
  tab_header(
    title = md("**Positive Pauls**"),
    subtitle = md("The 10 most positive commenters | *Min 10 comments*")
  ) %>% 
  cols_label(
    avg_sent = "Average Sentiment",
    tot_comments = "Comments"
  ) %>% 
  data_color(
    columns = vars(avg_sent),
    colors = scales::col_numeric(
      palette = as.character(paletteer_d("ggsci::light-green_material")),
      domain = NULL
    )
  ) %>% 
  tab_options(
    # Remove the top border from the table
    table.border.top.width = 0
  )

div(pos_paul, class='gt-no-scroll')
```

### Negative Nigels {.no-title}
```{r "negative-nigels"}
neg_nige <- mt_comms[,
  .(tot_comments=.N, avg_sent=mean(google_score))
  ,by=author
][tot_comments>10][order(avg_sent)][1:10] %>% 
  gt(rowname_col="author", id='neg-nigels-tab') %>% 
  tab_header(
    title = md("**Negative Nigels**"),
    subtitle = md("The 10 most negative commenters | *Min 10 comments*")
  ) %>% 
  cols_label(
    avg_sent = "Average Sentiment",
    tot_comments = "Comments"
  ) %>% 
  data_color(
    columns = vars(avg_sent),
    colors = scales::col_numeric(
      palette = as.character(paletteer_d("ggsci::red_material", direction = -1)),
      domain = NULL
    )
  ) %>% 
  tab_options(
    table.border.top.width = 0
  )
div(neg_nige, class='gt-no-scroll')
```

### Frequent Freddies {.no-title}
```{r "frequent-freddies"}
freq_fred <- mt_comms[
  author != "[deleted]", 
  .(tot_comments=.N), 
  by=author
][order(-tot_comments)][1:10] %>% 
  gt(rowname_col="author", id='freq-freddies-tab') %>% 
  tab_header(
    title = md("**Frequent Freddies**"),
    subtitle = md("The 10 most frequent commenters")
  ) %>% 
  cols_label(
    tot_comments = "Total Comments"
  ) %>% 
  data_color(
    columns = vars(tot_comments),
    colors = scales::col_numeric(
      palette = as.character(paletteer_d("ggsci::light-blue_material")),
      domain = NULL
    )
  ) %>% 
  tab_options(
    table.border.top.width = 0
  )

div(freq_fred, class='gt-no-scroll')
```


Row
-----------------------------------------------------------------------
### Ref Sentiment {.no-title .plotly-chart data-width=2}

```{r "ref-sent-plot"}
ref_data <- combodt[,.(posix_dt, rm_2mins_google_ref, cum_google_ref, google_score, ref)]

ref_data[,ref := ifelse(ref, 'Ref Comments', 'General Comments')]

refplot <- plot_ly(
  data = ref_data, 
  x = ~posix_dt, 
  colors = c(tier_cols$dark[[1]], tier_cols$dark[[2]])
) %>%
  add_trace(
    y = ~cum_google_ref,
    color = ~ref,
    name = ~ref,
    type = 'scatter',
    mode = 'lines',
    hoverinfo = 'y',
    visible = F
  ) %>%
  add_trace(
    y = ~rm_2mins_google_ref,
    color = ~ref,
    name = ~ref,
    type = 'scatter',
    mode = 'lines',
    hoverinfo = 'y'
  ) %>% 
  add_trace(
    data = score_data,
    type = 'scatter',
    mode = 'markers',
    marker = list(
      symbol = 'diamond', 
      color = ~event_bg_col, 
      size = 10, 
      line = list(color = ~event_border_col, width=2)
    ),
    x = ~posix_dt, 
    y = 0,
    text = ~graph_lab,
    hoverinfo = "text",
    hoverlabel = list(font = list(color = ~event_text_col), bordercolor = ~event_border_col)
  ) %>%
  layout(
    title = list(text = 'Referee Sentiment: 2 Minute Rolling Average'),
    yaxis = list(title = 'Average Sentiment'),
    xaxis = list(
      title = "Time",
      range = c(posix_KO, posix_FT),
      showspikes = T,
      spikemode = 'across',
      spikethickness = 2,
      spikecolor = 'black'
    ),
    hovermode = 'x',
    hoverdistance = 1,
    modebar = list(orientation = 'v'),
    legend = list(
      itemclick = F, 
      itemdoubleclick = F,
      y = -0.3,
      x = 0.5,
      xanchor = 'center',
      orientation = 'h'
    ),
    updatemenus = list(
      list(
        type = 'buttons',
        label = 'Category',
        y = -0.2,
        x = 0.5,
        xanchor = 'center',
        direction = 'left',
        buttons = list(
          list(
            method = "update",
            args = list(
              list('visible' = c(FALSE, FALSE, TRUE, TRUE)),
              list('yaxis.title' = 'Average Sentiment')
            ),
            label = "2 Min Rolling Avg"
          ),
          list(
            method = "update",
            args = list(
              list(visible=c(TRUE, TRUE, FALSE, FALSE)),
              list('yaxis.title' = 'Cumulative Sentiment')
            ),
            label = "Cumulative")
        )
      )
    )
  ) %>% 
  config(
    responsive = TRUE,
    modeBarButtonsToRemove = c(
      'hoverCompareCartesian', 
      'hoverClosestCartesian', 
      'lasso2d', 
      'select2d', 
      'toggleSpikelines'
    )
  )

refplot
```

### Ref Comment Stats {.no-title data-width=1}

```{r "ref-stats"}
ref_stats <- transpose(
  combodt[,.(
    total_comments = .N, 
    average_sentiment = mean(google_score), 
    negative_comment_pct = (sum(google_score < -0.25)/.N), 
    average_reddit_score = mean(score_praw)
    ), by=ref][, ref := ifelse(ref, 'Ref', 'General')][],
  keep.names = 'rn', 
  make.names = 'ref'
)

# Format rownames because gt doesn't :(
ref_stats[, rn := titleize(gsub('_', ' ', rn))]

ref_stats %>% 
  gt(rowname_col = 'rn') %>% 
  fmt(rows = 1, fns = as.numeric) %>% 
  fmt_number(columns = everything(), rows = 1, decimals = 0) %>%
  fmt_percent(columns = everything(), rows = 3, decimals = 1) %>% 
  fmt(rows = c(2, 4), fns = function(x){ sigfig(x, 3) })  %>% 
  tab_options(
    row.striping.include_stub = TRUE,
    row.striping.include_table_body = TRUE,
    table.border.top.width = 0
  ) %>% 
  tab_header(
    title = md("**Ref Comments v General Comments**")
  )
```

#### Ref by Fanbase {.header-central}
```{r "ref-fanbase"}
ref_rating <- combodt[
  ref == T, 
  .(rating = as.numeric(sigfig(mean(google_score)))),
  by=allegiance
]

ref_rating_html <- function(team_name, sentiment, team_cols, logos_pth) {
  team_0_sent_css = get_sentiment_col(sentiment) 
  return(list(
    div(class = 'ref-fanbase-logo-wrapper', 
      img(class = "ref-fanbase-logo", src=get_team_logo(team_name, team_cols, logos_pth)),
      p(
        titleize(team_name), 
        style = css(
          "padding" = "0px 5px",
          "border-width" = "1px", 
          "border-style" = "solid", 
          "border-radius" = "5px",
          "border-color" = get_team_border_col(team_name, team_cols),
          "color" = get_team_text_col(team_name, team_cols),
          "background-color" = get_team_bg_col(team_name, team_cols)
        )
      )
    ),
    div(class = 'ref-fanbase-rating-wrapper',
      p(class = glue("ref-fanbase-rating text-{team_0_sent_css} bg-{team_0_sent_css}"),
        sentiment,
        span(class = get_sentiment_icon(sentiment)),
        style = css(
          "border-width" = "1px", 
          "border-style" = "solid", 
          "border-radius" = "5px",
          "border-color" = glue("var(--{team_0_sent_css})")
        )
      )
    )
  ))
}


div(class = 'ref-fanbase-grid',
  ref_rating_html(TEAM_0_NAME, ref_rating[allegiance==TEAM_0_NAME, rating], team_cols, logos_pth),
  ref_rating_html('neutral', ref_rating[allegiance=='neutral', rating], team_cols, logos_pth),
  ref_rating_html(TEAM_1_NAME, ref_rating[allegiance==TEAM_1_NAME, rating], team_cols, logos_pth)
)

```

Row
-----------------------------------------------------------------------
### Emojis {.no-title .plotly-chart}
```{r "emojis"}
emojis <- unlist(sapply(mt_comms[['plaintext']], emo::ji_extract_all, USE.NAMES = F))
emojis <- data.table(table(emojis))
setorder(emojis, -N)
emojis[1:20, scaled := (N/min(N))/10]
emojis[as.data.table(emo::jis), on=c(emojis="emoji"), desc := i.name]
plot_ly(
  data = emojis, 
  x = ~N, y = ~emojis,
  text = ~desc, 
  hovertemplate = '<span style="text-shadow: 0px 0px 4px black">%{y}</span> | %{text} | Used: %{x} <extra></extra>'
) %>% 
  add_bars(marker = list(color = hex_to_rgba("#FFCC22")), hoverlabel = list(font = list(size = 17))) %>% 
  layout(
    title = list(text = 'Emoji Popularity'),
    yaxis = list(
      title = list(text = NULL),
      tickfont = list(size = 25),
      ticklen = 5,
      tickcolor = 'white',
      categoryorder = "total ascending", range = c(nrow(emojis)-10.5, nrow(emojis)-0.5)
    ),
    xaxis = list(title = 'Times Used'),
    modebar = list(orientation = 'v')
  ) %>% 
  config(
    responsive = TRUE,
    modeBarButtonsToRemove = c(
      'hoverCompareCartesian', 
      'hoverClosestCartesian', 
      'lasso2d', 
      'select2d', 
      'toggleSpikelines'
    )
  )
```

Row
-----------------------------------------------------------------------
```{r "swear-jar"}
mt_comms[,swears := strsplit(swears, ",")]
mt_comms[,swears_root := strsplit(swears_root, ",")]
author_swears_tab <- mt_comms[, 
  {n_swears=sum(sapply(swears, length))
   n_words=sum(words) 
   .(n_swears=n_swears, n_words=n_words, pct_swears=(n_swears/n_words)) },
  by=author
][n_words > 10,]
tot_swears_tab <- mt_comms[,.(swear=unlist(titleize(unlist(swears_root))))][, .(used=.N), by=swear][order(-used)]
```

### Total Swears {.no-title}
```{r "tot-swears"}
tot_swears_gt <- tot_swears_tab[1:10] %>% 
  gt(rowname_col="swear", id='tot-swears-tab') %>% 
  tab_header(
    title = md("**Most Common Swear Words**"),
    subtitle = md("The 10 most frequent swear words by root")
  ) %>% 
  cols_label(
    used = "Times Used"
  ) %>% 
  data_color(
    columns = vars(used),
    colors = scales::col_numeric(
      palette = as.character(paletteer_d("ggsci::orange_material")),
      domain = NULL
    )
  ) %>% 
  tab_options(
    table.border.top.width = 0
  )

div(tot_swears_gt, class='gt-no-scroll')
```

> The swear words covered and their "roots" can be found in <a href="https://github.com/awgymer/pyrugby/blob/master/reddit_profanities.json">this json</a>

### Author Swears Pct {.no-title}
```{r "author-swears-pct"}
author_swears_pct_gt <- author_swears_tab[order(-pct_swears)][1:10] %>% 
  gt(rowname_col="author", id='author-pct-swears-tab') %>% 
    tab_header(
      title = md("**Sweariest Commenters by Percent**"),
      subtitle = md("The 10 commenters with highest % of words as swear words | Min 10 words")
    ) %>% 
    fmt_percent(
      columns = vars(pct_swears),
      decimals = 2
    ) %>% 
    cols_label(
      n_swears = "Total Swears",
      n_words = "Total Words",
      pct_swears = "Swear Percentage"
    ) %>% 
    data_color(
      columns = vars(n_swears, n_words, pct_swears),
      colors = scales::col_numeric(
        palette = as.character(paletteer_d("ggsci::orange_material")),
        domain = NULL
      )
    ) %>% 
    tab_options(
      table.border.top.width = 0
    )

div(author_swears_pct_gt, class='gt-no-scroll')
```

### Author Swears Num {.no-title}
```{r "author-swears-num"}
author_swears_num_gt <- author_swears_tab[order(-n_swears)][1:10] %>% 
  gt(rowname_col="author", id='author-num-swears-tab') %>% 
    tab_header(
      title = md("**Sweariest Commenters by Volume**"),
      subtitle = md("The 10 commenters with highest number of swear words | Min 10 words")
    ) %>% 
    fmt_percent(
      columns = vars(pct_swears),
      decimals = 2
    ) %>% 
    cols_label(
      n_swears = "Total Swears",
      n_words = "Total Words",
      pct_swears = "Swear Percentage"
    ) %>% 
    data_color(
      columns = vars(n_swears, n_words, pct_swears),
      colors = scales::col_numeric(
        palette = as.character(paletteer_d("ggsci::orange_material")),
        domain = NULL
      )
    ) %>% 
    tab_options(
      table.border.top.width = 0
    )

div(author_swears_num_gt, class='gt-no-scroll')
```

